---
title: "Supplementary Material"
subtitle: "Tree diversity effects on soil ecosystem multifunctionality vary with proximity to target trees"
date: "Henriette Christel, Georg J.A. Hähn, Rémy Beugnon, Yuanyuan Huang, Olga Ferlian, Hafeez Ul Haq, Tesfaye Wubet, Nico Eisenhauer and Simone Cesarz"
output:
  pdf_document:
    fig_width: 15
    fig_height: 15
    fig_caption: yes
    toc: yes
---
```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE)

library("tidyverse")
library("readxl")
library("lmerTest")
library("optimx")
library("performance")
library("emmeans")
library("corrplot")
library("Hmisc")
library("reshape2")
library("patchwork")
library("car")
library("sjPlot")
library("ggpubr")
library("multcompView")
library("devtools")
library("multifunc")
library("gridExtra")
library("kableExtra")
library("knitr")
library("viridis")


MyDiv_data <- read_xlsx("../../02.data/MyDiv_TreeDi_data.xlsx")
MyDiv_2015 <- read_xlsx("../../02.data/pH-2015-dataportal/28_2_data.xlsx")

MyDiv_data <- MyDiv_data %>% 
  mutate(scale = case_when(
    type == "Rhizo" ~ "Target tree zone",
    type == "Bulk" ~ "Interaction zone")) %>% 
  mutate(dist = case_when(
    dist == 20 ~ 0, 
    dist == 70 ~ 1))

jitter <- position_jitter(width = 0.1)

```

## Figure S1: Correlation matrix


```{r, warning=FALSE, message=FALSE}
MyDiv_matrix <- MyDiv_data %>%
  select_if(is.numeric)
MyDiv_matrix <- MyDiv_matrix %>% 
  dplyr::select(cmic, basal, bgluc, xyl, nag, phos, WSA_perc, 
                mean_basal_area, C_content, N_content, pH, h2o_perc) 

MyDiv.corr <- cor(MyDiv_matrix, use = 'all') 

# plot
colnames(MyDiv.corr) = c(" CMIC", " BASAL", "BGLUC", "XYL", "NAG" ," PHOS",  
                         " WSA", " BA", " C", " N", " pH", " WATER")

rownames(MyDiv.corr) = c(" CMIC", " BASAL", "BGLUC", "XYL", "NAG" ," PHOS",  
                         " WSA", " BA", " C", " N", " pH", " WATER")

coloured.corr <- corrplot(MyDiv.corr, method="color",  
                          type = "lower", order = "original", 
                          addCoef.col = "black", # Add coefficient of correlation
                          tl.col = "black", tl.srt = 45, number.cex = 2.3, tl.cex = 2.5, cl.cex = 2, #Text label color and rotation
                          col = COL2('PuOr', 10), #color scheme 
                          sig.level = 0.01, insig = "blank", # hide correlation coefficient on the principal diagonal
                          diag = FALSE)
```

Figure S1: Correlation matrix of the measured single soil functions creating the multifunctionality index: microbial biomass (CMIC), basal respiration (BASAL), enzymatic activities of beta-glucosidase (BGLUC), beta-xylosidase (XYL), N-acetylglucosaminidase (NAG) and acid phosphatase (PHOS), mean basal area (BA) as a biotic predictor and the abiotic predictors soil carbon content (C), soil nitrogen content (N), soil pH (pH) and soil water content (WATER). Brownish colours indicate negative correlations, and purple indicates positive correlations. Due to the correlation of soil carbon and nitrogen content above 0.7, only soil carbon was used as an explanatory variable and nitrogen was excluded.

\newpage

## Table S2: Linear-mixed effects model outputs

###
Table S2: Linear-mixed effects model outputs of an analyses of variance (ANOVA) type I with F-values, degrees of freedom (NumDF, DenDF), p-values and effect directions. Plot mycorrhizal type (MYC) as arbuscular mycorrhizal fungi only (A), ectomycorrhizal fungi only (E) and a mixture of both (AE). Tree species richness (TSR) as monocultures, 2-, or 4-species mixtures. Sampling location at the small-scale (ZONE) as target tree zone (TZ) close to the target tree or interaction zone (IZ) in the centre between the target tree and the direct neighbours. Responses as: soil multifunctionality (MFI), microbial biomass (CMIC in µg microbial carbon g^-1^ dry weight), soil basal respiration (BASAL in µl O2 g^-1^ dry weight h^-1^), the enzymatic activity (in nmol cleaved substrate g^-1^ dry weight h^-1^) of beta-glucosidase (BGLUC), beta-xylosidase (XYL), N-acetylglucosaminidase (NAG) and acid phosphatase (PHOS), water-stable aggregates (WSA in %), soil pH (pH), soil water content (H~2~O in %), soil carbon content (C in %), and mean basal area (BA).

```{r, warning=FALSE, message=FALSE}
resp <- c("meanFunction", "cmic", "basal", "bgluc", "xyl", "nag", "phos", "WSA_perc", "pH", "h2o_perc", "C_content",
          "mean_basal_area")

for(i in 1:length(resp)) {
  if(resp[i] %in% c("cmic", "basal")) {
    mod <- lmer(as.formula(paste0(resp[i],"~ plot_myc + sp_nr + scale + 
                     plot_myc:sp_nr + 
                     plot_myc:scale + 
                     sp_nr:scale +
                     (1|block/plot) + (1|device)")), MyDiv_data)
  }
  
  if(resp[i] == "h2o_perc") {
    mod <- lmer(as.formula(paste0(resp[i],"~ plot_myc + sp_nr + scale + 
                     plot_myc:sp_nr + 
                     plot_myc:scale + 
                     sp_nr:scale +
                     (1|block/plot) + (1|sampling_date)")), MyDiv_data)
  }
  
  if(!(resp[i] %in% c("cmic", "basal", "h2o_perc"))) {
    mod <- lmer(as.formula(paste0(resp[i],"~ plot_myc + sp_nr + scale + 
                     plot_myc:sp_nr + 
                     plot_myc:scale + 
                     sp_nr:scale +
                     (1|block/plot)")), MyDiv_data)
    
  }
  
  anova <- anova(mod, type = 1) %>% as.data.frame()
  summary(mod)
  anova$response <- resp[i]
  anova$explanatory <- rownames(anova)
  
  if(i == 1) {
    anova.df <- anova
  } else{
    anova.df <- rbind(anova.df, anova)
  }
  
  
}

tab <- anova.df %>%
  mutate(Direction = c("A<AE<E", "+", "na", "na", "na", "TZ: +", 
                       "A<AE<E", "+", "na", "E: +", "na", "na",
                       "na", "na", "na", "na", "na","na",
                       "E<AE<A", "na", "na", "na", "na", "TZ: +",
                       "E<AE<A", "na", "na", "na", "na", "na",
                       "A<AE<E", "+", "na", "na", "na", "na",
                       "na", "+", "na", "na", "na", "na",
                       "A<AE<E", "na", "na", "na", "na", "na",
                       "na", "na", "-", "na", "na", "na",
                       "na", "-", "na", "na", "na", "na",
                       "na", "na", "na", "na", "na", "na",
                       "na", "na", "na", "na", "na", "na")) %>% 
  mutate(value = paste0("F: ", ifelse(round(`F value`, digits = 3)<0.001, "<0.001", round(`F value`, digits = 3)), "\n",
                        "DF: ", ifelse(round(`NumDF`, digits = 2)<0.001, "<0.001", round(`NumDF`, digits = 2)), ", " ,
                        ifelse(round(`DenDF`, digits = 2)<0.001, "<0.001", round(`DenDF`, digits = 2)), "\n",
                        "p: ", ifelse(round(`Pr(>F)`, digits = 3)<0.001, "<0.001", round(`Pr(>F)`, digits = 3)),
                        "\nDirect.: ", Direction)) %>%
  pivot_wider(id_cols = c(response), names_from = explanatory, values_from = value) %>%
  as.data.frame() %>% 
  mutate_all(linebreak) %>% 
  mutate(response = str_to_upper(response)) %>% 
  mutate(response = gsub("MEANFUNCTION", "MFI", response),
         response = gsub("WSA_PERC", "WSA", response),
         response = gsub("PH$", "pH", response),
         response = gsub("H2O_PERC", "$H_2O$", response),
         response = gsub("C_CONTENT", "C", response),
         response = gsub("MEAN_BASAL_AREA", "BA", response))
colnames(tab) <- c("Response", "MYC", "TSR", "ZONE", "TSR:MYC", "ZONE:MYC", "ZONE:TSR")
kable(tab %>% 
        dplyr::select(Response, TSR, MYC, `TSR:MYC`, ZONE, `ZONE:TSR`, `ZONE:MYC`), "latex", escape = FALSE)
```

\newpage

## Figure S2: Effects of tree species identity, sampling location and their interaction effects on soil multifunctionality

###

```{r, warning=FALSE, message=FALSE, include = FALSE}
MFI_tree_ID_AM <- MyDiv_data %>% filter(tree_myc == "AMF") %>% 
  ggplot() +
  geom_point(aes(x = tree_sp, y = meanFunction, col = type, shape = type), size = 5, alpha = 0.4, position = jitter) +
  geom_boxplot(aes(x = tree_sp, y = meanFunction, col = type, fill = type), alpha = 0.2) +
  theme_bw() +
  theme(legend.position = "top", 
        legend.text = element_text(size = 28),
        legend.title = element_text(size = 30),
        axis.title.x = element_text(size = 30), 
        axis.title.y = element_text(size = 30), 
        axis.text = element_text(size = 28),
        plot.title = element_text(hjust = 0.5, size = 26)) +
  labs(x= "AM tree species identity",  
       y = "MFI") +
  scale_color_manual(labels = c("Interaction zone", "Target tree zone"),
                     values = c("#298c8c", "#800074"),
                     name = "") +
  scale_fill_manual(labels = c("Interaction zone", "Target tree zone"),
                    values = c("#298c8c", "#800074"),
                    name = "") +
  scale_shape_manual(labels = c("Interaction zone", "Target tree zone"),
                     values = c(16, 17),
                     name = "") +
  scale_size_manual(labels = c("Interaction zone", "Target tree zone"),
                    values = c(5, 5),
                    name = "")

MFI_tree_ID_EM <- MyDiv_data %>% filter(tree_myc == "EMF") %>% 
    ggplot() +
    geom_point(aes(x = tree_sp, y = meanFunction, col = type, shape = type), size = 5, alpha = 0.4, position = jitter) +
    geom_boxplot(aes(x = tree_sp, y = meanFunction, col = type, fill = type), alpha = 0.2) +
    theme_bw() +
    theme(legend.position = "top", 
        legend.text = element_text(size = 28),
        legend.title = element_text(size = 30),
        axis.title.x = element_text(size = 30), 
        axis.title.y = element_text(size = 30), 
        axis.text = element_text(size = 28),
        plot.title = element_text(hjust = 0.5, size = 28)) +
    labs(x= "EM tree species identity",  
         y = "MFI") +
    scale_color_manual(labels = c("Interaction zone", "Target tree zone"),
                     values = c("#298c8c", "#800074"),
                     name = "") +
    scale_fill_manual(labels = c("Interaction zone", "Target tree zone"),
                      values = c("#298c8c", "#800074"),
                      name = "") +
    scale_shape_manual(labels = c("Interaction zone", "Target tree zone"),
                       values = c(16, 17),
                       name = "") +
  scale_size_manual(labels = c("Interaction zone", "Target tree zone"),
                    values = c(5, 5),
                    name = "")

supp.fig <- MFI_tree_ID_AM + MFI_tree_ID_EM + 
  plot_layout(nrow = 2) +
  plot_annotation(tag_levels = "a") +
  plot_layout(guides = "collect", axis_title = "collect") &
  theme(legend.position='top')
```

```{r}
supp.fig
```

Figure S2: Effects of tree species identity, sampling location (Interaction zone - teal, Target tree zone - magenta) and their interaction effects on soil multifunctionality (MFI). Tree species associated with arbuscular mycorrhizal fungi (AM, Panel a): Ac – *Acer pseudoplatanus*, Ae – *Aesculus hippocastanum*, Fr – *Fraxinus excelsior*, Pr – *Prunus arium*, So – *Sorbus aucuparia*. Tree species associated with ectomycorrhizal fungi (EM, Panel b): Be – *Betula pendula*, Ca – *Carpinus betulus*, Fa – *Fagus sylvatica*, Qu – *Quercus petraea*, Ti – *Tilia platyphyllos*.

\newpage

## Figure S3: Tree species richness, plot mycorrhizal type and their interaction effects on plot conditions

###

```{r, warning=FALSE, message=FALSE, include = FALSE}
pH_myc_div <- MyDiv_data %>% 
    ggplot() +
    geom_point(aes(x = sp_nr, y = pH, col = plot_myc, shape = plot_myc), size = 5, alpha = 0.4, position = jitter) +
    geom_smooth(aes(x = sp_nr, y = pH, col = plot_myc, fill = plot_myc), 
                method = "lm", alpha = 0.2) +   
    geom_smooth(aes(x = sp_nr, y = pH), method = "lm", colour = "#333333", alpha = 0, linewidth = 0.5) + 
    theme_bw() +
    scale_x_log10(breaks = c(1,2,4)) +
      theme(legend.position = "none", 
            legend.text = element_text(size = 28),
            legend.title = element_text(size = 30),
          axis.title.x = element_text(size = 30), 
          axis.title.y = element_blank(),  
          axis.text = element_text(size = 28),   
          plot.title = element_text(hjust = 0.5, size = 28)) +
    labs(x= "Tree species richness",  
         y = "", 
         title = "pH") +
    scale_color_manual(labels = c("AM", "MIX", "EM"), 
                       values = c("#93C47D", "#FFB000", "#6D9EEB"), 
                       name = "MYC:") +
    scale_fill_manual(labels = c("AM", "MIX", "EM"), 
                      values = c("#93C47D", "#FFB000", "#6D9EEB"), 
                      name = "MYC:") +
    scale_shape_manual(labels = c("AM", "MIX", "EM"), 
                       values = c(16, 15, 17),
                       name = "MYC:")


h2o_myc_div <- MyDiv_data %>% 
    ggplot() +
    geom_point(aes(x = sp_nr, y = h2o_perc, col = plot_myc, shape = plot_myc), size = 5, alpha = 0.4, position = jitter) +
    geom_smooth(aes(x = sp_nr, y = h2o_perc, col = plot_myc, fill = plot_myc), 
                method = "lm", alpha = 0.2) +   
    geom_smooth(aes(x = sp_nr, y = h2o_perc), method = "lm", colour = "#333333", alpha = 0, linewidth = 0.5) + 
    theme_bw() +
    scale_x_log10(breaks = c(1,2,4)) +
    theme(legend.position = "none", 
          legend.text = element_text(size = 28),
          legend.title = element_text(size = 30),
          axis.title.x = element_text(size = 30), 
          axis.title.y = element_text(size = 30),  
          axis.text = element_text(size = 28),        
          plot.title = element_text(hjust = 0.5, size = 28)) +
    labs(x= "Tree species richness",  
         y = "Value", 
         title = expression(H[2]*O)) +
    scale_color_manual(labels = c("AM", "MIX", "EM"), 
                       values = c("#93C47D", "#FFB000", "#6D9EEB"), 
                       name = "MYC:") +
    scale_fill_manual(labels = c("AM", "MIX", "EM"), 
                      values = c("#93C47D", "#FFB000", "#6D9EEB"), 
                      name = "MYC:") +
    scale_shape_manual(labels = c("AM", "MIX", "EM"), 
                       values = c(16, 15, 17),
                       name = "MYC:") +
    annotate(geom = "text", x = 1.1, y = 18, label = "TSR: ** ", size = 28/.pt)


carbon_myc_div <- MyDiv_data %>% 
    ggplot() +
    geom_point(aes(x = sp_nr, y = C_content, col = plot_myc, shape = plot_myc), size = 5, alpha = 0.4, position = jitter) +
    geom_smooth(aes(x = sp_nr, y = C_content, col = plot_myc, fill = plot_myc), 
                method = "lm", alpha = 0.2) +   
    geom_smooth(aes(x = sp_nr, y = C_content), method = "lm", colour = "#333333", alpha = 0, linewidth = 0.5) + 
    theme_bw() +
    scale_x_log10(breaks = c(1,2,4)) +
    theme(legend.position = "none", 
          legend.text = element_text(size = 28),
          legend.title = element_text(size = 30),
          axis.title.x = element_blank(), 
          axis.title.y = element_text(size = 30),  
          axis.text = element_text(size = 28),    
          plot.title = element_text(hjust = 0.5, size = 28)) +
    labs(x= "Tree species richness",  
         y = "Value", 
         title = "C") +
    scale_color_manual(labels = c("AM", "MIX", "EM"), 
                       values = c("#93C47D", "#FFB000", "#6D9EEB"), 
                       name = "MYC:") +
    scale_fill_manual(labels = c("AM", "MIX", "EM"), 
                      values = c("#93C47D", "#FFB000", "#6D9EEB"), 
                      name = "MYC:") +
    scale_shape_manual(labels = c("AM", "MIX", "EM"), 
                       values = c(16, 15, 17),
                       name = "MYC:") 

nitrogen_myc_div <- MyDiv_data %>% 
    ggplot() +
    geom_point(aes(x = sp_nr, y = N_content, col = plot_myc, shape = plot_myc), size = 5, alpha = 0.4, position = jitter) +
    geom_smooth(aes(x = sp_nr, y = N_content, col = plot_myc, fill = plot_myc), 
                method = "lm", alpha = 0.2) +   
    geom_smooth(aes(x = sp_nr, y = N_content), method = "lm", colour = "#333333", alpha = 0, linewidth = 0.5) + 
    theme_bw() +
    scale_x_log10(breaks = c(1,2,4)) +
    theme(legend.position = "none", 
          legend.text = element_text(size = 28),
          legend.title = element_text(size = 30),
          axis.title.x = element_blank(), 
          axis.title.y = element_blank(),  
          axis.text = element_text(size = 28),    
          plot.title = element_text(hjust = 0.5, size = 28)) +
    labs(x= "Tree species richness",  
         y = "", 
         title = "N") +
    scale_color_manual(labels = c("AM", "MIX", "EM"), 
                       values = c("#93C47D", "#FFB000", "#6D9EEB"), 
                       name = "MYC:") +
    scale_fill_manual(labels = c("AM", "MIX", "EM"), 
                      values = c("#93C47D", "#FFB000", "#6D9EEB"), 
                      name = "MYC:") +
    scale_shape_manual(labels = c("AM", "MIX", "EM"), 
                       values = c(16, 15, 17),
                       name = "MYC:") 

MyDiv_data$CN <- MyDiv_data$C_content/MyDiv_data$N_content
cn_myc_div <- MyDiv_data %>% 
    ggplot() +
    geom_point(aes(x = sp_nr, y = CN, col = plot_myc, shape = plot_myc), size = 5, alpha = 0.4, position = jitter) +
    geom_smooth(aes(x = sp_nr, y = CN, col = plot_myc, fill = plot_myc), 
                method = "lm", alpha = 0.2) +   
    geom_smooth(aes(x = sp_nr, y = CN), method = "lm", colour = "#333333", alpha = 0, linewidth = 0.5) + 
    theme_bw() +
    scale_x_log10(breaks = c(1,2,4)) +
    theme(legend.position = "none", 
          legend.text = element_text(size = 28),
          legend.title = element_text(size = 30),
          axis.title.x = element_blank(), 
          axis.title.y = element_blank(),  
          axis.text = element_text(size = 28),          
          plot.title = element_text(hjust = 0.5, size = 28)) +
    labs(x= "Tree species richness",  
         y = "", 
         title = "C/N") +
    scale_color_manual(labels = c("AM", "MIX", "EM"), 
                       values = c("#93C47D", "#FFB000", "#6D9EEB"), 
                       name = "MYC:") +
    scale_fill_manual(labels = c("AM", "MIX", "EM"), 
                      values = c("#93C47D", "#FFB000", "#6D9EEB"), 
                      name = "MYC:") +
    scale_shape_manual(labels = c("AM", "MIX", "EM"), 
                       values = c(16, 15, 17),
                       name = "MYC:") 

layt = "AABBCC"

pt <-  carbon_myc_div + nitrogen_myc_div + cn_myc_div + 
    plot_annotation(tag_levels = "a") +
    plot_layout(guides = "collect", axis_titles = "collect_x") &
    theme(legend.position='none')

pb <- h2o_myc_div + pH_myc_div +
    plot_annotation(tag_levels = list(c("d", "e"))) +
    plot_layout(guides = "collect", axis_titles = "collect_x") &
    theme(legend.position='bottom')

abiotic <- pt / pb + plot_annotation(tag_levels = "a")
```

```{r, message=FALSE}
abiotic
```
Figure S3: Tree species richness (TSR), plot mycorrhizal type (MYC, AM = arbuscular mycorrhizal fungi only - green, EM = ectomycorrhizal fungi only - blue, Mix = mixture of AM and EM - yellow) and their interaction effects on soil carbon content (C in %), soil nitrogen content (N in %), carbon to nitrogen ratio (C/N), soil pH and soil water content (H~2~O in %). The results of linear-mixed effects models for significant relationships are provided on the top left of each panel (* p<.05; ** p<.01; *** p<.001).

\newpage

## Figure S4: Tree species richness, sampling location and their interaction effects on plot conditions

###

```{r, warning=FALSE, message=FALSE, include = FALSE}
pH_div_type <- MyDiv_data %>% 
    ggplot() +
    geom_point(aes(x = sp_nr, y = pH, col = type, shape = type), size = 5, alpha = 0.4, position = jitter) +
    geom_smooth(aes(x = sp_nr, y = pH, col = type, fill = type), 
                method = "lm", alpha = 0.2) +
    theme_bw() +
    scale_x_log10(breaks = c(1,2,4)) +
    theme(legend.position = "none",  
          legend.text = element_text(size = 28),
          legend.title = element_text(size = 30),
          axis.title.x = element_text(size = 30), 
          axis.title.y = element_blank(), 
          axis.text = element_text(size = 28),
          plot.title = element_text(hjust = 0.5, size = 28)) +
    labs(x= "Tree species richness",  
         y = "", 
         title = "pH") +
    scale_color_manual(labels = c("Interaction zone", "Target tree zone"), 
                       values = c("#298c8c", "#800074"),
                       name = "") +
    scale_fill_manual(labels = c("Interaction zone", "Target tree zone"), 
                      values = c("#298c8c", "#800074"),
                      name = "") +
    scale_shape_manual(labels = c("Interaction zone", "Target tree zone"), 
                       values = c(16, 17),
                       name = "") +
    annotate(geom = "text", x = 1.18, y = 7.45, label = "ZONE: ***", size = 28/.pt)

h2o_div_type <- MyDiv_data %>% 
    ggplot() +
    geom_point(aes(x = sp_nr, y = h2o_perc, col = type, shape = type), size = 5, alpha = 0.4, position = jitter) +
    geom_smooth(aes(x = sp_nr, y = h2o_perc, col = type, fill = type), 
                method = "lm", alpha = 0.2) +
    theme_bw() +
    scale_x_log10(breaks = c(1,2,4)) +
    theme(legend.position = "none",  
          legend.text = element_text(size = 28),
          legend.title = element_text(size = 30),
          axis.title.x = element_text(size = 30), 
          axis.title.y = element_text(size = 30), 
          axis.text = element_text(size = 28),
          plot.title = element_text(hjust = 0.5, size = 28)) +
    labs(x= "Tree species richness",  
         y = "Value", 
         title = expression(H[2]*O)) +
    scale_color_manual(labels = c("Interaction zone", "Target tree zone"), 
                       values = c("#298c8c", "#800074"),
                       name = "") +
    scale_fill_manual(labels = c("Interaction zone", "Target tree zone"), 
                      values = c("#298c8c", "#800074"),
                      name = "") +
    scale_shape_manual(labels = c("Interaction zone", "Target tree zone"), 
                       values = c(16, 17),
                       name = "")

carbon_div_type <- MyDiv_data %>% 
    ggplot() +
    geom_point(aes(x = sp_nr, y = C_content, col = type, shape = type), size = 5, alpha = 0.4, position = jitter) +
    geom_smooth(aes(x = sp_nr, y = C_content, col = type, fill = type), 
                method = "lm", alpha = 0.2) +
    theme_bw() +
    scale_x_log10(breaks = c(1,2,4)) +
    theme(legend.position = "none",  
          legend.text = element_text(size = 28),
          legend.title = element_text(size = 30),
          axis.title.x = element_blank(), 
          axis.title.y = element_text(size = 30), 
          axis.text = element_text(size = 28),
          plot.title = element_text(hjust = 0.5, size = 28)) +
    labs(x= "Tree species richness",  
         y = "Value", 
         title = "C") +
    scale_color_manual(labels = c("Interaction zone", "Target tree zone"), 
                       values = c("#298c8c", "#800074"),
                       name = "") +
    scale_fill_manual(labels = c("Interaction zone", "Target tree zone"), 
                      values = c("#298c8c", "#800074"),
                      name = "") +
    scale_shape_manual(labels = c("Interaction zone", "Target tree zone"), 
                       values = c(16, 17),
                       name = "")

nitrogen_div_type <- MyDiv_data %>% 
    ggplot() +
    geom_point(aes(x = sp_nr, y = N_content, col = type, shape = type), size = 5, alpha = 0.4, position = jitter) +
    geom_smooth(aes(x = sp_nr, y = N_content, col = type, fill = type), 
                method = "lm", alpha = 0.2) +
    theme_bw() +
    scale_x_log10(breaks = c(1,2,4)) +
    theme(legend.position = "none",  
          legend.text = element_text(size = 28),
          legend.title = element_text(size = 30),
          axis.title.x = element_blank(), 
          axis.title.y = element_blank(), 
          axis.text = element_text(size = 28),
  plot.title = element_text(hjust = 0.5, size = 28)) +
    labs(x= "Tree species richness",  
         y = "", 
         title = "N") +
    scale_color_manual(labels = c("Interaction zone", "Target tree zone"), 
                       values = c("#298c8c", "#800074"),
                       name = "") +
    scale_fill_manual(labels = c("Interaction zone", "Target tree zone"), 
                      values = c("#298c8c", "#800074"),
                      name = "") +
    scale_shape_manual(labels = c("Interaction zone", "Target tree zone"), 
                       values = c(16, 17),
                       name = "")

MyDiv_data$CN <- MyDiv_data$C_content/MyDiv_data$N_content
cn_div_type <- MyDiv_data %>% 
    ggplot() +
    geom_point(aes(x = sp_nr, y = CN, col = type, shape = type), size = 5, alpha = 0.4, position = jitter) +
    geom_smooth(aes(x = sp_nr, y = CN, col = type, fill = type), 
                method = "lm", alpha = 0.2) +
    theme_bw() +
    scale_x_log10(breaks = c(1,2,4)) +
    theme(legend.position = "none",  
          legend.text = element_text(size = 28),
          legend.title = element_text(size = 30),
          axis.title.x = element_blank(), 
          axis.title.y = element_blank(), 
          axis.text = element_text(size = 28),
          plot.title = element_text(hjust = 0.5, size = 28)) +
    labs(x= "Tree species richness",  
         y = "", 
         title = "C/N") +
    scale_color_manual(labels = c("Interaction zone", "Target tree zone"), 
                       values = c("#298c8c", "#800074"),
                       name = "") +
    scale_fill_manual(labels = c("Interaction zone", "Target tree zone"), 
                      values = c("#298c8c", "#800074"),
                      name = "") +
    scale_shape_manual(labels = c("Interaction zone", "Target tree zone"), 
                       values = c(16, 17),
                       name = "")


layt = "AABBCC"

pt <-  carbon_div_type + nitrogen_div_type + cn_div_type + 
  plot_annotation(tag_levels = "a") +
  plot_layout(guides = "collect", design = layt, axis_titles = "collect_x") &
  theme(legend.position='none')

pb <- h2o_div_type + pH_div_type +
  plot_annotation(tag_levels = list(c("d", "e"))) +
  plot_layout(guides = "collect", axis_titles = "collect_x") &
  theme(legend.position='bottom')

abiotic <- pt / pb + plot_annotation(tag_levels = "a")
```


```{r, message = FALSE}
abiotic
```

Figure S4: Tree species richness (TSR), sampling location (ZONE, Interaction zone - teal, Target tree zone - magenta) and their interaction effects on soil carbon content (C in %), soil nitrogen content (N in %), carbon to nitrogen ratio (C/N), soil pH and soil water content (H~2~O in %). The results of linear-mixed effects models for significant relationships are provided on the top left of each panel (* p<.05; ** p<.01; *** p<.001).

\newpage

## Figure S6: Activity ratio of carbon to nitrogen cycle-related enzymes

###

```{r, warning=FALSE, message=FALSE, include = FALSE}
MyDiv_data$bn <- MyDiv_data$bgluc/MyDiv_data$nag

bn.mod <- lmer(bn ~ plot_myc + (1|block/plot), MyDiv_data %>% filter(MyDiv_data$plot_myc != "AE"))
anova(bn.mod, type = 1)

emmeans(bn.mod, list(pairwise ~ plot_myc), adjust = "tukey")

bgluc_nag <- MyDiv_data %>% 
    filter(MyDiv_data$plot_myc != "AE") %>% 
    #mutate(scale = factor(scale, levels = c("Plot level", "Target tree zone", "Interaction zone"))) %>% 
    ggplot() +
    geom_point(aes(x = plot_myc, y = bn, col = bn), size = 5, alpha = 0.4, position = jitter) +
    geom_boxplot(aes(x = plot_myc, y = bn, col = bn), alpha = 0.2) +
    geom_line(data = data.frame(x = c(1, 2),
                                y = c(11.5, 11.5)), aes(x = x, y = y)) +
    geom_line(data = data.frame(x = c(1, 1),
                                y = c(11.5, 11)), aes(x = x, y = y)) +
    geom_line(data = data.frame(x = c(2, 2),
                                y = c(11.5, 11)), aes(x = x, y = y)) +
    theme_bw() +
    theme(legend.text = element_text(size = 28),
          legend.title = element_text(size = 30),
          axis.text.x = element_text(size = 28), 
          axis.text.y = element_text(size = 28),
          axis.title.x = element_text(size = 30),
          axis.title.y = element_text(size = 30)) +
    labs(x = "Plot mycorrhizal type", 
         y = "BGLUC / NAG ratio") +
    scale_color_viridis(name = "activity ratio", 
                        breaks = c(2, 5.5, 9)) +
    annotate(geom = "text", x = 1.5, y = 11.8 , label = "***", size = 28/.pt)
```

```{r, warning=FALSE}
bgluc_nag
```

Figure S6: Activity ratio of carbon to nitrogen cycle-related enzymes. beta-glucosidase (BGLUC) active in the carbon cycle and N-acetylglucosaminidase (NAG) active in the nitrogen cycle. Yellow indicates a high ratio with high activity of BGLUC and low activity of NAG, purple indicates a low ratio with low activity of BGLUC and high activity of NAG. The results of the Tukey-test are provided above the boxes (* p<.05; ** p<.01; *** p<.001).

\newpage 

## Figure S7: Comparison of the soil pH in the year of establishment (2015) and year of sampling (2021)

###

```{r, warning=FALSE, message=FALSE, include = FALSE}

df2015 <- MyDiv_2015 %>% 
  select(year, plotName, soil_pH) %>% 
  rename(plotname = plotName, 
         pH = soil_pH) %>% 
  mutate(scale = "Plot level")

df2021 <- MyDiv_data %>% 
  select(plotname, pH, scale) %>% 
  mutate(year = 2021)

MyDiv <- df2015 %>% rbind(df2021)

mod.pH <- lmer(pH ~ scale + (1|plotname), MyDiv)
summary(mod.pH)
anova(mod.pH, type = 1)

emmeans(mod.pH, list(pairwise ~ scale), adjust = "tukey")


pH_year <- MyDiv %>% 
    mutate(scale = factor(scale, levels = c("Plot level", "Target tree zone", "Interaction zone"))) %>% 
    ggplot() +
    geom_point(aes(x = year, y = pH, col = scale, shape = scale), size = 5, alpha = 0.4, position = jitter) +
    geom_boxplot(aes(x = year, y = pH, col = scale, fill = scale), alpha = 0.2) +
    geom_line(data = data.frame(x = c(1, 1.81),
                                y = c(7.97, 7.97)), aes(x = x, y = y)) +
    geom_line(data = data.frame(x = c(1, 1),
                                y = c(7.97, 7.93)), aes(x = x, y = y)) +
    geom_line(data = data.frame(x = c(1.81, 1.81),
                                y = c(7.97, 7.93)), aes(x = x, y = y)) +
    geom_line(data = data.frame(x = c(1, 2.19),
                                y = c(7.87, 7.87)), aes(x = x, y = y)) +
    geom_line(data = data.frame(x = c(1, 1),
                                y = c(7.87, 7.83)), aes(x = x, y = y)) +
    geom_line(data = data.frame(x = c(2.19, 2.19),
                                y = c(7.87, 7.83)), aes(x = x, y = y)) +
    geom_line(data = data.frame(x = c(1.81, 2.19),
                                y = c(7.57, 7.57)), aes(x = x, y = y)) +
    geom_line(data = data.frame(x = c(1.81, 1.81),
                                y = c(7.57, 7.53)), aes(x = x, y = y)) +
    geom_line(data = data.frame(x = c(2.19, 2.19),
                                y = c(7.57, 7.53)), aes(x = x, y = y)) +
    theme_bw() +
    theme(legend.position = "top", 
          legend.text = element_text(size = 28),
          legend.title = element_text(size = 30),
          axis.text.x = element_text(size = 28),
          axis.text.y = element_text(size = 28),
          axis.title.x = element_blank(),
          axis.title.y = element_text(size = 30)) +
    scale_color_manual(values = c("#54457f", "#800074", "#298c8c"),
                       name = "") +
    scale_fill_manual(values = c("#54457f", "#800074", "#298c8c"),
                      name = "") +
    scale_shape_manual(values = c(16, 17, 18),
                       name = "") +
    annotate(geom = "text", x = c(1.405, 1.595, 2), y = c(8, 7.9, 7.6) , label = c("***", "***", "***"), size = 28/.pt)
```

```{r}
pH_year
```

Figure S7: Comparison of the soil pH in the year of establishment (2015) and year of sampling (2021). Violet colour shows the plot-level data from 2015, the data of 2021 is divided in the two sampling locations close to the target tree (Target tree zone - magenta) and in the centre between the target tree and the direct neighbours (Interaction zone - teal). The results of the Tukey-test are provided above the boxes (* p<.05; ** p<.01; *** p<.001).
